---
title: "Homework 3"
author: Sanjana Batabyal 
output: github_document
---

This is my solution to Homework 3.

```{r}
library(tidyverse)
library(magrittr)
library(p8105.datasets)
```

## Problem 1
Loading the data.
```{r}
data("instacart")
```

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
There are 134 aisles and the aisles that are most ordered from are fresh vegetables (with 150,609 items) and fresh fruits (with 150,473 items).

Making a plot that shows the number of items ordered in each aisle.
```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x=aisle, y=n)) +
  geom_point()
```
Making a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetable fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n)))%>%
  filter(rank<4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```
Making a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
